Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-07-18 paper.tex
Comments on intro and section II

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{minted} % Used for JSON highlighting
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}

@@

Although Linux Containers (LXC) and other alternatives such as Rocket were already
available, container-based virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. Docker makes container's use more straightforward. However, there
are few limitations of Docker. For example, it is not supported by shared High Performance
Computing (HPC) environments because its default configuration runs containers
as root. Secondly, Docker containers cannot be used for reproducibility as Docker
images can be updated, and hence losing specific package versions required for
reproducing scientific results. Consequently, Singularity was introduced in 2016 with the main
goal of removing drawbacks of Docker. Singularity was developed mainly to
support HPC environments and to provide reproducibility.

No doubt containers have gained a lot of popularity after Docker, however containers tightly
integrate with the host due to the sharing of the kernel, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments.
The goal of this paper is to report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis.

@@

@@

In this section we summarize a particular use case of containers, Canadian
Open Neuroscience Platform (CONP). It is a national platform for sharing neuroscience
research data and aims at creating an interactive interface for neuroscientists and
clinical neuroscience. The main goal of the CONP platform is to store, process, and
distribute massive amounts of data produced by modern neuroscience.
Here, we summarize all the resources that are
used by CONP to function.

@@

Compute Canada is a national High Performance Computing (HPC) research platform.
It provides infrastructure and services for Advanced Research Computing (ARC).
Its goal is to bring excellence, ease, and innovation in research of Canadian
research institutions by providing storage, computing power, and required
softwares to carry out research.
It integrates with four regional partners namely ACENET, Calcul Québec, Compute
Ontario, and WestGrid for providing local support.
Compute Canada uses Slurm Workload Manager to schedule jobs and Lustre shared
file system to provide high bandwidth to its users.
Control groups are used by it to limit number of resources that can be used
by a particular user.
Compute Canada do not support Docker as its default configuration
runs containers as root. However, Compute Canada supports Singularity
containers because it removes the drawback of Docker.

@@

@@

A container image is a package that contains the application alongwith its
required dependencies and libraries.
Nowadays, two most popular containerization tools used are Docker and Singularity.
Both have different image format. Below we provide description of both image
formats.

@@

Docker images are organized as a series of layers which are stacked on top
of each other. Each layer consists of a filesystem diff that is introduced
due to the changes made on the layer below it. All layers of the Docker image
are identified by unique layerIDs. All these layers stacked together gives a
unified and complete view of a Docker image. These layers are
compressed into a single image and can be pushed to the Docker Hub.
There are two ways to create Docker image. The first is to use an existing
Docker image as a base image and then build a new image from that one, by
making changes in the filesystem, which is saved as a new layer of the
Docker image. The second way is to build a new Docker image from scratch.

@@

Singularity takes snapshot, locks, and archives the developed application, which
is known as Singularity image. Due to image hashing, Singularity does not
contain image layers, unlike Docker images. So Singularity image is a single
immutable file that can be moved around as any other file. While publishing experimental
results, authors can also publish Singularity image along with its hash, which
allow other researchers to verify results.

@@

Following are the application frameworks in neuroscience that are using containers
on HPC clusters.

@@

Boutiques is a application sharing system~\cite{glatard2015boutiques}. It is based on Linux containers
and can be used to publish, integrate, and execute third-party
applications across different platforms automatically. Boutiques framework is
employed by CONP to integrate and execute applications on CBRAIN web portal.
With Boutiques, the command-line of the application is
described through a flexible template, which is called descriptor. Descriptor
describes the input that is required by the application and the output it produces~\cite{glatard2018boutiques}.
It points to a Docker or Singularity image, where the application is installed.
Developers of the application produce Boutiques descriptor, and stores it along
with the application. Boutiques descriptor are described in JSON format. There
is a set of core tools which help in constructing, validating and executing bou-
tiques descriptors. During the application run-time, the execution platform
builds a command line by using input values given by the user and boutiques
descriptor of the application.
Currently, Boutiques tools use 23 different container images. Out of
which 7 are Singularity images and rest are Docker images.

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different operating system users as well as HPC users through containerization.
BIDS apps share reproducibility goal with Boutiques, though they are conceptually different.
BIDS apps main goal is to standardize application interfaces whereas Boutiques goal is to
make application sharing and integration easier with different platforms.
Currently, there are 27 different container images that are used by
BIDS App. All these are Docker images.

@@

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images.

@@

There are various scanning tools that are available for scanning container
images like Vuls, Clair, ClairScanner, Anchore, Dagda, etc.
Out of these we take Anchore, Vuls, and ClairScanner tools for our experiments to scan container
images which are used by the above discussed application frameworks.
For scanning Singularity images, Stools scanner is used, which is specifically designed
for Singularity images. Stools internally make use of Clair-Scanner for
scanning.

@@

Anchore is an end-to-end container security, and open-source compliance plat-
form. Anchore is a static scanning tool that analyses container images and
lists operating system packages,
unofficial packages, configuration files, and language modules [58]. In other
words, Anchore is a set of tools, which users can employ for scanning container
images and applying custom policies to them. It provides visibility and trans-
parency of the container environment. Anchore offers services in two ways:
Anchore Engine and Anchore Engine Command Line Interface (CLI).
Anchore Engine is available as a Docker image, which is available on Docker
Hub. Anchore Engine CLI provides command line interface in addition to An-
chore Engine REST API. Anchore Engine CLI can be installed directly on the
host and used to inspect images.

@@

Vuls is an open-source vulnerability scanner for Linux/FreeBSD, written in Go
language. It is an agentless vulnerability scanner and uses various
vulnerability databases. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to de-
tect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server. As shown in Figure 4.3
it is clear that vulnerability detection time of Vuls is much better than An-
chore. Vuls tool uses dynamic scanning, so the scanning time does not depend
on the image size.
Graph 4.5 shows percentage of vulnerabilities that are
covered by Vuls and Anchore tools. This graph can be explained with the help
of List 4.4, which shows vulnerability databases that are referred, by both Vuls
and Anchore. Anchore covers less percentage of vulnerabilities as compared
to Vuls because Anchore relies mainly on vulnerability databases published by
Linux distributions, whereas Vuls additionally refers to another sources such as
databases from Japan, US-CERT, etc.

@@

Stools combines CoreOS's ClairScanner with the Continuous Integration (Travis and
Circle) for scanning Singularity images for vulnerabilities.
Initially, ClairScanner was intented to scan Docker Containers. This tool does not fit well
to scan Singularity images because of the different image formats. Docker layers
consist of layers, whereas Singluarity images consist of a single binary file.
However, TravisCI and CircleCI are commonly used for testing code, and these
services also offer running containers. So Stools combined both the concepts.
It involves spinning a ClairScanner container during testing and building a
Singularity image, and scanning it for filesystem before the image is finalized.

>>>>

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}

@@

Although Linux Containers (LXC) and other alternatives such as Rocket were already
available, container-based virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. Docker makes container's use more straightforward. However, there
are few limitations of Docker. For example, it is not supported by shared High Performance
Computing (HPC) environments because its default configuration runs containers
as root. Secondly, Docker containers cannot be used for reproducibility as Docker
images can be updated, and hence losing specific package versions required for
reproducing scientific results. Consequently, Singularity was introduced in 2016 with the main
goal of removing drawbacks of Docker. Singularity was developed mainly to
support HPC environments and to provide reproducibility. \TG{The previous two paragraphs are too general. You should
summarize them in 2-3 sharper sentences.}

No doubt containers have gained a lot of popularity after Docker, however containers tightly
integrate with the host due to the sharing of the kernel \TG{not only the kernel: 
mounted file systems and devices too}, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

\TG{Avoid bold face in papers}

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments.
The goal of this paper is to report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. \TG{You should explain the 3 questions discussed during out meeting.}

@@

\TG{Some notions that should be introduced in this section (incomplete):
vulnerability vs malware, paper suggested by Emad Shihab.}

@@

In this section we summarize a particular use case of containers, Canadian
Open Neuroscience Platform (CONP) \TG{No. In this section you describe the data and tools 
that were used in your experiments. The fact that they are used in CONP 
is a side note}. It is a national platform for sharing neuroscience
research data and aims at creating an interactive interface for neuroscientists and
clinical neuroscience. The main goal of the CONP platform is to store, process, and
distribute massive amounts of data produced by modern neuroscience.
Here, we summarize all the resources that are
used by CONP to function.

@@

Compute Canada is a national High Performance Computing (HPC) research platform.
It provides infrastructure and services for Advanced Research Computing (ARC).
Its goal is to bring excellence, ease, and innovation in research of Canadian
research institutions by providing storage, computing power, and required
softwares to carry out research.
It integrates with four regional partners namely ACENET, Calcul Québec, Compute
Ontario, and WestGrid for providing local support.
Compute Canada uses Slurm Workload Manager to schedule jobs and Lustre shared
file system to provide high bandwidth to its users.
Control groups are used by it to limit number of resources that can be used
by a particular user.
Compute Canada do not support Docker as its default configuration
runs containers as root. However, Compute Canada supports Singularity
containers because it removes the drawback of Docker.
\TG{Not sure if Compute Canada is used in the experiments. If not, remove and 
pitch it in the introduction that containers are nowadays used in many clusters, including 
Compute Canada. You could also refer to the full list of clusters that support singularity.}

@@

\TG{This paragraph is too general for a method section. You might summarize this in
2-3 sentences for the introduction. Here you should talk about the engine,
i.e., how images are run, focusing on security features. Use of namespaces,
cgroups, access to devices, etc}

\TG{You should add a subsection presenting Docker.}

@@

A container image is a package that contains the application along with its
required dependencies and libraries.
Nowadays, two most popular containerization tools used are Docker and Singularity.
Both have different image format. Below we provide description of both image
formats.

@@

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers \TG{If a layer is reused
in different images, for instance the top layer of a base image, will it
have the same id across images?}. Each layer consists of a filesystem diff
with respect to the layer below. Layers are compressed \TG{Are you
referring to file compression similar to gzip?} into a single image and can
be pushed to DockerHub. There are two ways to create a Docker image. The
first is to use an existing Docker image as a base image and then build a
new image from that one, by making changes in the filesystem, which is
saved as a new layer of the Docker image. The second way is to build a new
Docker image from scratch. \TG{It would be relevant to mention the file
system used to store the images (is it squashfs?).}

@@

Singularity takes snapshot, locks, and archives the developed application
\TG{what does it mean?}, which is known as Singularity image. Due to image
hashing, Singularity does not contain image layers, unlike Docker images
\TG{is it really due to hashing? why couldn't a hashed image have layers?}.
So Singularity image is a single immutable file that can be moved around as
any other file \TG{Talking about the file system used would be relevant}.
While publishing experimental results, authors can also publish Singularity
image along with its hash, which allow other researchers to verify results.

@@

We used containerized applications available in the Boutiques and BIDS Apps
frameworks used in neuroscience to run analyses on HPC clusters.

@@

Boutiques~\cite{glatard2018boutiques} is a application sharing system to
publish, integrate, and execute command-line applications across different
platforms. In Boutiques, the command-line of the application is described
through a flexible JSON descriptor pointing to a Docker or a Singularity
image where the application is installed. A set of core tools help
construct, validate, publish and execute Boutiques descriptors. Boutiques
applications are published on the Zenodo research repository.

We used the 49 Boutiques applications published on Zenodo at the time of
the writing. These applications use 23 different container images,
including 7 Singularity images and 16 Docker images. \TG{You should add a
table showing the images available in each framework, with their base
image, DOIs, size, and number of files.}

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
\TG{How many BIDS apps are there? Did you use them all?} Currently, there are 27 different container images that are used by
BIDS App. All these are Docker images.

@@

\subsection{Other images}

\TG{We could also use selected images from DockerHub or SingularityHub. For
instances the ones used by AFNI or other neuroimaging tools not represented
in Boutiques or BIDS Apps. However we wouldn't be able to shrink these
images unless we identified specific command-lines to run.}

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images. \TG{Here you should describe the method you used to update the images, linking to the 
software (script) you used to do that. 
This paragraph is too general and doesn't belong to the methods. }

@@

Out of the various container image scanning tools, we selected Anchore,
Vuls, Stools and ClairScanner for our experiments. \TG{Explain why you used
more than 1 tool, and why you used these ones speifically.}

@@

Anchore is an end-to-end, open-source container security platform. Anchore
is a static \TG{You should explain static vs dynamic scanning in the intro}
scanning tool that analyses container images and lists operating system
packages, unofficial packages, configuration files \TG{How about files
which are not configuration files?}, and language modules [58] \TG{Import
reference}. \TG{What is a language module?} \TG{Does it list unknown
packages, files or modules?} Anchore is a set of tools, which users can
employ for scanning container images and applying custom policies to them.
It provides visibility and transparency of the container environment.
Anchore offers services in two ways: Anchore Engine and Anchore Engine
Command Line Interface (CLI). Anchore Engine is available as a Docker
image, which is available on Docker Hub. Anchore Engine CLI provides
command line interface in addition to Anchore Engine REST API. Anchore
Engine CLI can be installed directly on the host and used to inspect
images.

\TG{You should refer to the vulnerability databases used, and explain how
software components, for instance files, are matched to these databases. Is
the matching based on file checksum?}

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD, written in Go
language \TG{Why is it relevant that it's written in Go?}.
 It is an agentless vulnerability scanner and uses various
vulnerability databases \TG{You should list them and explain how they differ to 
the ones used by Anchore}. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to detect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server. As shown in Figure 4.3 \TG{There is no figure 4.3 or graph 4.5 here. Anyway it's not relevant 
to report results in this section. This paragraph looks like a copy paste from your previous report, it needs to be edited.}
it is clear that vulnerability detection time of Vuls is much better than An-
chore. Vuls tool uses dynamic scanning, so the scanning time does not depend
on the image size.
Graph 4.5 shows percentage of vulnerabilities that are
covered by Vuls and Anchore tools. This graph can be explained with the help
of List 4.4, which shows vulnerability databases that are referred, by both Vuls
and Anchore. Anchore covers less percentage of vulnerabilities as compared
to Vuls because Anchore relies mainly on vulnerability databases published by
Linux distributions, whereas Vuls additionally refers to another sources such as
databases from Japan, US-CERT, etc.

@@

Stools combines CoreOS's ClairScanner with the Continuous Integration (Travis and
Circle) for scanning Singularity images for vulnerabilities.
Initially, ClairScanner was intented to scan Docker Containers. This tool does not fit well
to scan Singularity images because of the different image formats. Docker layers
consist of layers, whereas Singluarity images consist of a single binary file.
However, TravisCI and CircleCI are commonly used for testing code, and these
services also offer running containers. So Stools combined both the concepts.
It involves spinning a ClairScanner container during testing and building a
Singularity image, and scanning it for filesystem before the image is finalized.
\TG{Add link and/or reference.}
